Pyg4ometry: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-01 public/paper/pyg4ometry.tex
Paper update

\section{GDML to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 
solids can be mapped to a FLUKA \verb|REGION|. Some solids in Geant4 directly map to 
a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between geant4 solids and the bodies used to compose a fluka region.

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/CuboidalChamber.pdf}
\caption{Example conversion of a simple GDML geometry to FLUKA, left panel :  
the original model in pyg4ometry, right panel : the converted FLUKA geometry 
viewed in pyg4ometry. The example is a sector bend dipole electromagnet.}
\label{fig:flair-gdml}
\end{center}
\end{figure*}

@@

@@

\subsection{Disjunctive normal form and degenerate surfaces}
Typically FLUKA will decompose a \verb|REGION| into disjunctive normal form (DNF), this normal form is 
characterised as the union of intersections, so 
\begin{equation}
R = z_1 \; | \;z_2\;  | \; z_3 	\; | \; z_4 \dots
\end{equation}
Defining \verb|REGION|s in terms of DNF allows the rapid test if a point is inside or outside. Testing each zone 
$z_i$ of $R$ can terminate if a point is determined to be inside any of convex zones. In general 
Equation~\ref{eqn:logicalSubtraction} is not in a DNF. If there are many levels of logical-physical volume 
placement, then recursive application of  Equation~\ref{eqn:logicalSubtraction} will create a nested set of 
parentheses. There are some conditions where a general boolean expression can yield an exponential 
explosion of the final DNF. There are well known algorithms to convert logical expressions to DNF 
\colorbox{yellow}{XXX}. Pyg4ometery can simplify parentheses from a \verb|REGION| by creating a 
sympy \cite{10.7717/peerj-cs.103} expression and using its \verb|logical| module. Before version 4.0 FLUKA 
would simplify all regions to DNF and some converted geometry would generate too many zones and would fail.
FLUKA version 4 allows control of the DNF expansion.

>>>>

\section{GDML to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 
solids can be mapped to a FLUKA \verb|REGION|. Some solids in Geant4 directly map to 
a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between Geant4 solids and the bodies used to compose a FLUKA region.

@@

Figure~\ref{fig:flair-gdml} shows an example conversion from GDML to FLUKA. The example is a 
vacuum chamber with three CF beam pipes connected to CF flanges. The top and bottom plates 
have been removed to better see the geometry. The model is formed of \verb|G4Box| and
\verb|G4Tubs| and boolean operations of subtraction, intersection and union.
\begin{figure*}[htb!]
\begin{center}
\includegraphics[width=1.0\textwidth]{./model-scene/CuboidalChamber.pdf}
\caption{Example conversion of a simple GDML geometry to FLUKA, left panel :  
the original model in pyg4ometry, right panel : the converted FLUKA geometry 
viewed in pyg4ometry. The example is a sector bend dipole electromagnet.}
\label{fig:flair-gdml}
\end{center}
\end{figure*}

@@

\subsection{Disjunctive normal form and degenerate surfaces}
Typically FLUKA will decompose a \verb|REGION| into disjunctive normal form (DNF), this normal form is 
characterised as the union of intersections, so 
\begin{equation}
R = z_1 \; | \;z_2\;  | \; z_3 	\; | \; z_4 \dots
\end{equation}
Defining \verb|REGION|s in terms of DNF allows the rapid test if a point is inside or outside. Testing each zone 
$z_i$ of $R$ can terminate if a point is determined to be inside any of convex zones. In general 
Equation~\ref{eqn:logicalSubtraction} is not in a DNF. If there are many levels of logical-physical volume 
placement, then recursive application of  Equation~\ref{eqn:logicalSubtraction} will create a nested set of 
parentheses. There are some conditions where a general boolean expression can yield an exponential 
explosion of the final DNF. There are well known algorithms to convert logical expressions to DNF 
\colorbox{yellow}{XXX}. Pyg4ometery can simplify parentheses from a \verb|REGION| by creating a 
sympy \cite{10.7717/peerj-cs.103} expression and using its \verb|logical| module \verb|to_ndf| method based on 
the Quine–McCluskey algorithm \cite{6769983}. Before version 4.0 FLUKA  would simplify all regions to 
DNF and some converted geometry would generate too many zones and would fail. FLUKA version 4 
allows control of the DNF expansion.
